---
title: "CANOE FRRF report"
output: pdf_document
date: "2023-07-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(plot3D)
require(FRRF)
require(oce)
require(marelac)

source("FRRFscripts.R")
```

# Underway data

Time in the underway data is in CEST, which is two hours later than the greenland time, on which our clocks were set.

```{r, echo=FALSE}
load("../../cruise/processed_data/underway.rda")
load("../../cruise/processed_data/events.rda")
load("../../cruise/accessory/LATLONcoast.rda")
load("../../cruise/processed_data/stations.rda") 
J0 <- 19553
Greenland <- subset(UnderWay, subset=longitude <= -42)
Greenland$Timestamp <- as.POSIXct(Greenland$Timestamp)
```

# FRRF station data

```{r}
load(file = "../processed_data/FRRFlist.rda")     # FRRFlist
load(file = "../processed_data/FRRFunderway.rda") # FRRFunderway
```

## Standardize the FRRF data 

```{r}
frrflist <- FRRFlist  # Take a copy
FRRFlist <- NULL
ff <- unique(frrflist$file)
for (i in ff){
  frrf <- subset(frrflist, subset=file==i)
  FRRFlist <- rbind(FRRFlist,
                  standardizeFRRF(frrf=frrf, 
                            Fblanc=frrf$Fblanc, 
                            keep = c("file", "depth", "repl", "JVPII_est",
                                      "latitude", "longitude")))
}

FRRFfiles <- unique(FRRFlist[,c("file", "station", "date", "depth", "repl", 
                                "latitude", "longitude", "Fblanc")])

```

```{r}
frrfunderway <- FRRFunderway
FRRFunderway <- NULL
ff <- unique(frrfunderway$file)
for (i in ff){
  frrf <- subset(frrfunderway, subset=file==i)
  FRRFunderway <- rbind(FRRFunderway,
                  standardizeFRRF(frrf=frrf, 
                            Fblanc=frrf$Fblanc, 
                            keep = c("file", "JVPII_est", 
                                     "longitude", "latitude")))
}

UNDERWAYfiles <- unique(FRRFunderway[,c("file", "station", "date", 
                                "latitude", "longitude", "Fblanc")])
```

## PE curve

The JVPII variable is fitted as a function of light. It is in units of 
mmol electrons m-3 h-1. 

To convert JVPII to mg C m3 h-1, we use the electron requirement for 
C-fixation. We use 5 (mol elektronen/mol C):

	P = JVPII/PHIe,C × 12

```{r, fig.width=8, fig.height=10, warning=FALSE}
par(mfrow=c(5,5), mar=c(4,4,4,0))
inot <- which(FRRFlist$Fo < 0) 

#response <- "JVPIIest"
response = "JVPII"

if (length(inot)) FRRFlist <- FRRFlist[-inot, ]
Uin <- unique(FRRFlist$file)
FIT <- NULL
for (f in Uin) {
  i <- which (FRRFsettings$station==f)
  Pmax <- FRRFsettings$Emax[i]
  frrf <- subset(FRRFlist, subset=file==f)
  station <- paste(frrf$station[1],FRRFsettings$depth[i], "m")
  FIT <- rbind(FIT, data.frame(fitFRRF(frrf=frrf, station=station, PARmax=Pmax, response=response), file=f, stat=frrf$station[1]))
}
names(FIT)[names(FIT)=="station"] <- "descriptor"
names(FIT)[names(FIT)== "stat"] <- "station"

```

# Resulting parameters

```{r}

FF <- merge(FIT, FRRFfiles, by="file")
names(FF)[names(FF)=="station.x"] <- c("station")
FF[ii <- which(is.infinite(FF$rsq)),]  # deep stations -> remove them
if (length(ii)) FF <- FF[-ii, ]

FF[ii <- which(FF$alpha>100),]  # deep stations -> remove them
if (length(ii)) FF <- FF[-ii, ]

FF[ii <- which(FF$nrow<4),]  # deep stations -> remove them
if (length(ii)) FF <- FF[-ii, ]

#FF[ii <- which(FF$ps>100),]  # deep stations -> remove them

names(FF)[names(FF)==c("station", "station.y")] <- c("file", "station")

knitr::kable(FF[,c("station", "depth", "repl", "alpha", "eopt", "ps", "meanChl")], 
             digits=c(0,0,0,4,0,2,2))
```

```{r, fig.width=8, fig.height=8}
FF$station <- factor(FF$station, 
                        levels = c("2O", "4O", "5O", "6B", "7F", "8O", "9B", 
                                   "10O", "11M", "12O", "13B", "14O", "15F", 
                                   "15O", "16O", "17O", "19O", "45O", "48F",
                                   "49O", "50B", "51O", "52B", "53B", "54O",
                                   "55O", "56F", "57O", "58B", "59B", "60B" ))
    
  
par(mfrow=c(1,3))
with (FF, boxplot(alpha~station, horizontal = TRUE, las=1, outline=FALSE))
with (FF, boxplot(ps~station, horizontal = TRUE, las=1, outline=FALSE))
with (FF, boxplot(meanChl~station, horizontal = TRUE, las=1, outline=FALSE))


```

## Photosynthesis

To convert JVPII to mg C m3 h-1, we use the electron requirement for 
C-fixation:

	P = JVPII/PHIe,C × 12
	
We assume PHIe,C = 5 (mol elektronen/mol C).The factor 12 is to go from mmol to mgram.

To estimate production, we need the lightintensity entering the water (E0 (µmol fotonen m 2 s 1 ) and the extinction coefficient (Kd m-1).

The latter has been estimated for each station by fitting the light versus depth profiles, as:

Iz = 

I0*(p * exp(-k1*z) + (1-p) * exp(-k2*z))

```{r}
Lightpar <- read.csv("../../ctd/processed_data/IrradianceFit.csv")
LightFun <- function(p=0.5, k1=0.1, k2=1, irr=1, 
                     D=seq(0, 100, length.out=100))
  irr*(p*exp(-k1*D)+(1-p)*exp(-k2*D))
```

The underway data hold the light intensity, in W/m2.

Light is converted from W/m2 to uEinst/m2/s, assuming 4.6 uE/m2/s per W/m2. We assume that 50% of Light is PAR.

```{r}
Light <- Greenland[, c("Timestamp", "WeatherStation..CR6.WS..1..XDR.Solar.radiation.density..W.m2.")]
colnames(Light)[1:2] <- c("time", "I_W.m2")
Light <- subset(Light, subset= time>"2023-07-17" & time <= "2023-07-22")
Light$par <- Light$I_W.m2*4.6*0.5

plot(Light$time, Light$par, type="l")
```

Primary production of a station

```{r}
stat <- "15F"
statnr <- 15

PPstat <- function(
    stat, statnr,        # to select PP and light parameters
    fac = 1/5*12*24,     # convert from mol e/m2/hr to mgC/m2/d
    DD  = seq(from=0, to=100, by=0.5)) # Depth layers (top, bottom)
  {  
  deltaD <- diff(DD)

  DD <- 0.5*(DD[-1] + DD[-length(DD)])  # mean depth

# photosynthesis parameters, averaged per depth 
  # select station
  Fsub <- subset(FF, 
                 subset=station==stat)[, c("depth", "alpha", "eopt", "ps")]
  
  # average of parameters at the depth
  F <- data.frame(alpha=tapply(Fsub$alpha, INDEX=Fsub$depth, FUN=mean),
                  eopt =tapply(Fsub$eopt , INDEX=Fsub$depth, FUN=mean),
                  ps   =tapply(Fsub$ps,    INDEX=Fsub$depth, FUN=mean))
  Fdepth <- as.integer(row.names(F))

  # each depth matched with these parameters - assume they are valid till 
  # the middle with the next depth
  fD <- c(0, 0.5*(Fdepth[-1]+Fdepth[-length(Fdepth)]), 1000)

#  P <- NULL
#  for (i in 1:length(DD))
#    P <- rbind(P, F[max(1, which(fD[-length(fD)]<=DD[i] & fD[-1] > DD[i])),])

  P <- cbind(approx(Fdepth, F$alpha, xout=DD, rule=2)$y,
             approx(Fdepth, F$eopt, xout=DD, rule=2)$y,
             approx(Fdepth, F$ps, xout=DD, rule=2)$y)
  
  # light extinction parameters for this station
  Isub <- colMeans(subset(Lightpar, subset=station==statnr)[,-(1:3)])

  Irel <- LightFun(p=Isub[1], k1=Isub[2], k2=Isub[3], irr=1, D=DD)

  # Function to estimate the production per layer; 
  EPprof <- function(p, PAR) # p and PAR are matrix/vector
    PAR/((1/(p[,1]*p[,2]^2))*PAR^2+(1/p[,3]-2/(p[,1]*p[,2]))*PAR+(1/p[,1]))

  Ptime <- function(irr)
    sum(EPprof(P, Irel*irr)*deltaD)

  PP <- sapply(Light$par, FUN=Ptime)
  list(total = PP*fac, profile = EPprof(P, Irel*100), depth=DD)
}

PP <- data.frame(time=Light$time, PAR=Light$par,
                 p2O=PPstat ("2O", 2)$total, 
                 p4O=PPstat ("4O", 4)$total,
                 p5O=PPstat ("5O", 5)$total, 
                 p6B=PPstat ("6B", 6)$total,
                 p7F=PPstat ("7F", 8)$total, 
                 p8O=PPstat ("8O", 8)$total,  # NOTE: NO LIGHT DATA FOR ST 7
                 p9B=PPstat ("9B", 9)$total,
                 p10O=PPstat("10O",10)$total, 
                 p11M=PPstat("11M",11)$total, 
                 p12O=PPstat("12O",12)$total, 
                 p13B=PPstat("13B",13)$total,
                 p14O=PPstat("14O",14)$total, 
                 p15F=PPstat("15F",15)$total, 
                 p15O=PPstat("15O",15)$total, 
                 p16O=PPstat("16O",16)$total,
                 p17O=PPstat("17O",17)$total, 
                 p19O=PPstat("19O",19)$total)
Prof <- data.frame(depth=PPstat ("2O", 2)$depth,
                 p2O=PPstat ("2O", 2)$profile, 
                 p4O=PPstat ("4O", 4)$profile,
                 p5O=PPstat ("5O", 5)$profile, 
                 p6B=PPstat ("6B", 6)$profile,
                 p7F=PPstat ("7F", 8)$profile, 
                 p8O=PPstat ("8O", 8)$profile,  # NOTE: NO LIGHT DATA FOR ST 7
                 p9B=PPstat ("9B", 9)$profile,
                 p10O=PPstat("10O",10)$profile, 
                 p11M=PPstat("11M",11)$profile, 
                 p12O=PPstat("12O",12)$profile, 
                 p13B=PPstat("13B",13)$profile,
                 p14O=PPstat("14O",14)$profile, 
                 p15F=PPstat("15F",15)$profile, 
                 p15O=PPstat("15O",15)$profile, 
                 p16O=PPstat("16O",16)$profile,
                 p17O=PPstat("17O",17)$profile, 
                 p19O=PPstat("19O",19)$profile)


TOTPP <- data.frame(station=c(2,4,5,6,7,8,9,10,11,12,13,14,15,15,16,17,19),
                    prod=colMeans(PP[,-(1:2)]))
```

```{r}
Eudepth <- read.csv("eudepth.csv")
Fu     <- droplevels(subset(FF, subset=depth<50 & station != "52B"))
PIpars <- data.frame(station=c(2,4,5,6,7,8,9,10,11,12,13,14,15,15,16,17,19),
                     alpha= tapply(Fu$alpha, INDEX=Fu$station, FUN=mean),
                     ps=tapply(Fu$ps,        INDEX=Fu$station, FUN=mean),
                     eopt=tapply(Fu$eopt,    INDEX=Fu$station, FUN=mean),
                     chl=tapply(Fu$meanChl,  INDEX=Fu$station, FUN=mean),
                     euphotic=Eudepth$euphotic,
                     totPP = TOTPP$prod)
with(PIpars, plot(station, alpha))
PAR <- 0:500
plot(EP  (p = c(PIpars$alpha[1], eopt = PIpars$eopt[1], ps = 1), #PIpars$ps[1]), 
          PAR = PAR), type="l", lwd=2)
nr <- nrow(PIpars)
for (i in 2:nr)
lines(EP(p=c(PIpars$alpha[i], eopt = PIpars$eopt[i], ps = 1), #PIpars$ps[i]), 
          PAR = PAR), col=i)
```

```{r, fig.width=8, fig.height=10}
par(mfrow=c(3,1), las=1, cex.main=2, cex.lab=1.5, cex.axis=1.3)
plot(PIpars$station, PIpars$chl, xlab="station nr", ylab="ug/L", pch=16,
     type="b", main= "mean Chlororophyll conc", cex=2)
plot(PIpars$station, -PIpars$euphotic, xlab="station nr", ylab="m", pch=16,
     type="b", main= "euphotic depth", ylim=c(-25,0), cex=2, col="blue")
plot(PIpars$station, PIpars$totPP, xlab="station nr", ylab="mgC/m2/d", pch=16,
     type="b", main= "Photosynthesis", col="darkgreen", cex=2)
```







```{r, fig.width=8, fig.height=10, warning=FALSE}
par(mfrow=c(4,4), mar=c(4,4,4,0))
inot <- which(FRRFunderway$Fo < 0) 

#response <- "JVPIIest"
response = "JVPII"

if (length(inot)) FRRFunderway  <- FRRFunderway [-inot, ]
Uin <- unique(FRRFunderway$file)
FITunderway <- NULL
for (f in Uin) {
#  i <- which (FRRFsettings$station==f)
  Pmax <- 200
  frrf <- subset(FRRFunderway, subset=file==f)
  station <- paste(f, ":", format(frrf$latitude[1], digits=5), "N,",
                                  format(frrf$longitude[1], digits=5), "E")
  FITunderway <- rbind(FITunderway, 
                       data.frame(fitFRRF(frrf=frrf, station=station, 
                                          PARmax=Pmax, response=response), 
                                  file=f, stat=frrf$station[1]))
}
names(FITunderway)[names(FITunderway)=="station"] <- "descriptor"
names(FITunderway)[names(FITunderway)== "stat"] <- "station"

```
